---
title: "ShootingProject"
author: "S.Smiley"
date: "2024-08-03"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
# NYC Shooting Dataset Background
  This report covers an analysis of the NYC shooting dataset found at "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD".  The dataset lists every shooting incident that occurred in NYC going back to 2006 through the end of the previous calendar year.  This is a breakdown of every shooting incident that occurred in NYC going back to 2006 through the end of the previous calendar year. This data is manually extracted every quarter and reviewed by the Office of Management Analysis and Planning before being posted on the NYPD website. Each record represents a shooting incident in NYC and includes information about the event, the location and time of occurrence. In addition, information related to suspect and victim demographics is also included. This data can be used by the public to explore the nature of shooting/criminal activity. 

# Clear Statement of the Question of Interest
  The question of interest I have from this dataset is:
  **Does a victim's age, race, and sex indicate who the perpetrator might be?**
  
## Import Libraries
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
# Install necessary packages if they are not already installed
if (!requireNamespace("readr", quietly = TRUE)) {
  install.packages("readr")
}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
if (!requireNamespace("VIM", quietly = TRUE)) {
  install.packages("VIM")
}
if (!requireNamespace("nnet", quietly = TRUE)) {
  install.packages("nnet")
}
if (!requireNamespace("pwr", quietly = TRUE)) {
  install.packages("pwr")
}
# import libraries
library(readr)
library(dplyr)
library(ggplot2)
library(VIM)
library(tidyr)
library(nnet)
library(pwr)

```
# Data Loading
## Get Shooting Data
  Using the link of where the data comes from is a much more **reproducible** form of loading the data.
```{r get_shooting_data}
url_names <- c("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")

shooting_data <- read_csv(url_names[1])
head(shooting_data)
```
# Data Cleaning
## Convert "(null)" strings to NA
```{r fix_null}
shooting_data[shooting_data == "(null)"] <- NA
```
## Remove unnecessary columns
```{r remove_bad_cols}
columns_to_keep <- c("OCCUR_DATE", "BORO", "PRECINCT", "PERP_AGE_GROUP", "PERP_SEX", "PERP_RACE", "VIC_AGE_GROUP", "VIC_SEX", "VIC_RACE","STATISTICAL_MURDER_FLAG")
shooting_data <- shooting_data %>%
  select(all_of(columns_to_keep))
head(shooting_data)
```  
## Summary w/o cleaning
```{r show_sum_no_clean}
summary(shooting_data)
```
## Total up missing data in each column
```{r missing_data_cnt}
missing_counts <- colSums(is.na(shooting_data))
print(missing_counts)

```

## Visualize missing data
```{r viz_missing}
library(VIM)
aggr(shooting_data, col=c('green','red'), numbers=TRUE, sortVars=TRUE, labels=names(shooting_data), cex.axis=.5, gap=4, ylab=c("missing data","pattern"))
```

## Create an "UNKONWN" value for the missing data fields
  It appears a significant amount of people might have gotten away with murder since over 30% of the missing data is from the perpetrator.  Therefore, we don't want to omit this data.  Instead, we want to just note that it is "UNKNOWN."  This should help us have less bias in conclusions on shootings and murders since we would have to make some major assumptions otherwise.
```{r replace_missing}
clean_data <- shooting_data %>%
  mutate(
    PERP_AGE_GROUP = replace_na(PERP_AGE_GROUP, "UNKNOWN"),
    PERP_SEX = replace_na(PERP_SEX, "UNKNOWN"),
    PERP_RACE = replace_na(PERP_RACE, "UNKNOWN"),
    VIC_AGE_GROUP = replace_na(VIC_AGE_GROUP, "UNKNOWN"),
    VIC_SEX = replace_na(VIC_SEX, "UNKNOWN"),
    VIC_RACE = replace_na(VIC_RACE, "UNKNOWN")
  )
```
  
## Total up missing data in each column again
```{r missing_data_cnt_again}
missing_counts <- colSums(is.na(clean_data))
print(missing_counts)

```

## Convert date to Date format
```{r fix_date}
clean_data <- clean_data %>%
  mutate(OCCUR_DATE = as.Date(OCCUR_DATE, format = "%m/%d/%Y"))
head(clean_data)
```

## Convert character columns to factors
```{r fix_char}
clean_data <- clean_data %>%
  mutate(across(where(is.character), as.factor))
head(clean_data)
```

## Summary w cleaning
```{r show_sum_clean}
summary(clean_data)
```
# Data Analysis
## Bar plot of Perpetrator Age Group vs. Victim Age Group
```{r perp_vs_vic_age,include=TRUE}
ggplot(clean_data, aes(x = PERP_AGE_GROUP, fill = VIC_AGE_GROUP)) +
  geom_bar(position = "dodge") +
  labs(title = "Perpetrator Age Group vs. Victim Age Group",
       x = "Perpetrator Age Group",
       y = "Count",
       fill = "Victim Age Group") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) 
```

## Notice odd age groups
Odd Perpetrator Age groups: 1020, 940, 224, 1028
Odd Victim Age groups: 1022
Replace with UNKNOWN for specific odd age groups for perpetrators and victims
```{r remove_odd_age,include=TRUE}
odd_perp_age_groups <- c("1020", "940", "224", "1028")
odd_vic_age_groups <- c("1022")

clean_data <- clean_data %>%
  mutate(
    PERP_AGE_GROUP = case_when(
      PERP_AGE_GROUP %in% odd_perp_age_groups ~ "UNKNOWN",
      TRUE ~ PERP_AGE_GROUP
    ),
    VIC_AGE_GROUP = case_when(
      VIC_AGE_GROUP %in% odd_vic_age_groups ~ "UNKNOWN",
      TRUE ~ VIC_AGE_GROUP
    )
  )
clean_data <- clean_data %>%
  mutate(across(where(is.character), as.factor))
head(clean_data)
```
## Bar plot of Perpetrator Age Group vs. Victim Age Group - Verify Age 
Notice the victim and perpetrator age groups of 18-24 & 25-44 are the highest in these shooting of all known age groups.
Again, perpetrator's UNKNOWN is significant relative to the age, and it appears reasonable to assume the UNKNOWN age is similar to their victim's age from this chart.
```{r perp_vs_vic_age_verify,include=TRUE}
ggplot(clean_data, aes(x = PERP_AGE_GROUP, fill = VIC_AGE_GROUP)) +
  geom_bar(position = "dodge") +
  labs(title = "Perpetrator Age Group vs. Victim Age Group",
       x = "Perpetrator Age Group",
       y = "Count",
       fill = "Victim Age Group") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) 
```

## Bar plot by Perpetrator Sex
```{r perp_vs_vic_sex,include=TRUE}
ggplot(clean_data, aes(x = PERP_AGE_GROUP, fill = VIC_SEX)) +
  geom_bar(position = "dodge") +
  facet_wrap(~ PERP_SEX) +
  labs(title = "Perpetrator Age Group vs. Victim Sex Faceted by Perpetrator Sex",
       x = "Perpetrator Age Group",
       y = "Count",
       fill = "Victim Sex") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) 
```

## Replace sex "U" with "UNKNOWN"
```{r fix_U_UNKONWN_sex}
# add levels to sex for UKNOWN if not in it
levels(clean_data$PERP_SEX) <- c(levels(clean_data$PERP_SEX), "UNKNOWN")
levels(clean_data$PERP_AGE_GROUP) <- c(levels(clean_data$PERP_AGE_GROUP), "UNKNOWN")
levels(clean_data$PERP_RACE) <- c(levels(clean_data$PERP_RACE), "UNKNOWN")
levels(clean_data$VIC_SEX) <- c(levels(clean_data$VIC_SEX), "UNKNOWN")
levels(clean_data$VIC_AGE_GROUP) <- c(levels(clean_data$VIC_AGE_GROUP), "UNKNOWN")
levels(clean_data$VIC_RACE) <- c(levels(clean_data$VIC_RACE), "UNKNOWN")
clean_data <- clean_data %>%
  mutate(
    PERP_SEX = replace(PERP_SEX, PERP_SEX == "U", "UNKNOWN"),
    VIC_SEX = replace(VIC_SEX, VIC_SEX == "U", "UNKNOWN")
  )

```

## Bar plot by Perpetrator Sex Clean
```{r perp_vs_vic_sex_clean,include=TRUE}
ggplot(clean_data, aes(x = PERP_AGE_GROUP, fill = VIC_SEX)) +
  geom_bar(position = "dodge") +
  facet_wrap(~ PERP_SEX) +
  labs(title = "Perpetrator Age Group vs. Victim Sex Faceted by Perpetrator Sex",
       x = "Perpetrator Age Group",
       y = "Count",
       fill = "Victim Sex") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) 
```


## Bar plot by Perpetrator Race vs. Age Group
```{r perp_vs_vic_race,include=TRUE}
ggplot(clean_data, aes(x = PERP_AGE_GROUP, fill = VIC_RACE)) +
  geom_bar(position = "dodge") +
  facet_wrap(~ PERP_RACE) +
  labs(title = "Perpetrator Age Group vs. Victim Race Faceted by Perpetrator Race",
       x = "Perpetrator Age Group",
       y = "Count",
       fill = "Victim Race") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),axis.title = element_text(angle=45,size=8),strip.text = element_text(size = 5))
```

## Bar plot by Perpetrator Race vs. Sex
It appears that Black Males are the majority of shooters among all races and genders.
```{r perp_vs_vic_race_sex,include=TRUE}
ggplot(clean_data, aes(x = PERP_SEX, fill = VIC_RACE)) +
  geom_bar(position = "dodge") +
  facet_wrap(~ PERP_RACE) +
  labs(title = "Perpetrator SEX vs. Victim Race Faceted by Perpetrator Race",
       x = "Perpetrator SEX",
       y = "Count",
       fill = "Victim Race") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),axis.title = element_text(angle=45,size=8),strip.text = element_text(size = 5))
```

# Model Selection/training
## Multinomial logistic regression model
## Remove unused levels
When training a model, we need to ensure that the levels are consistent.
```{r clean_levels}
clean_data <- droplevels(clean_data)
```
```{r contin_tables}
# Contingency table for Victim Age Group vs. Perpetrator Age Group
table_vic_perp_age <- table(clean_data$VIC_AGE_GROUP, clean_data$PERP_AGE_GROUP)
print(table_vic_perp_age)

# Contingency table for Victim Race vs. Perpetrator Race
table_vic_perp_race <- table(clean_data$VIC_RACE, clean_data$PERP_RACE)
print(table_vic_perp_race)

# Contingency table for Victim Sex vs. Perpetrator Sex
table_vic_perp_sex <- table(clean_data$VIC_SEX, clean_data$PERP_SEX)
print(table_vic_perp_sex)

# Unique values for victim and perpetrator characteristics
unique_vic_race <- unique(clean_data$VIC_RACE)
unique_vic_sex <- unique(clean_data$VIC_SEX)
unique_vic_age_group <- unique(clean_data$VIC_AGE_GROUP)

unique_perp_race <- unique(clean_data$PERP_RACE)
unique_perp_sex <- unique(clean_data$PERP_SEX)
unique_perp_age_group <- unique(clean_data$PERP_AGE_GROUP)

# Generate all combinations of perpetrator race, gender, and age group
all_combinations <- expand.grid(
  VIC_RACE = unique_vic_race,
  VIC_SEX = unique_vic_sex,
  VIC_AGE_GROUP = unique_vic_age_group,
  PERP_RACE = unique_perp_race,
  PERP_SEX = unique_perp_sex,
  PERP_AGE_GROUP = unique_perp_age_group
)

# View the first few rows of the generated combinations
head(all_combinations)

```
## Model for PERP_AGE_GROUP
```{r age_model_perp}
age_model <- multinom(PERP_AGE_GROUP ~ VIC_AGE_GROUP + VIC_RACE + VIC_SEX, data = clean_data)
summary(age_model)
```

## Model for PERP_RACE
```{r race_model_perp}
race_model <- multinom(PERP_RACE ~ VIC_AGE_GROUP + VIC_RACE + VIC_SEX, data = clean_data)
summary(race_model)
```

## Model for PERP_SEX
```{r sex_model_perp}
sex_model <- multinom(PERP_SEX ~ VIC_AGE_GROUP + VIC_RACE + VIC_SEX, data = clean_data)
summary(sex_model)
```

## Inference with Model on Sample Data Point
Define a single data point for victim characteristics
```{r single_point_data}
# Define a single data point for victim characteristics
single_data_point <- clean_data[1, c("VIC_AGE_GROUP", "VIC_RACE", "VIC_SEX")]
single_data_point[1, ] <- list("25-44", "WHITE", "M")
```
### Predict the perpetrator's age group
```{r predict_age}
predicted_age_group <- predict(age_model, newdata = single_data_point)
print(paste("Predicted Perpetrator Age Group:", predicted_age_group))
```
### Predict the perpetrator's race
```{r predict_race}
predicted_race <- predict(race_model, newdata = single_data_point)
print(paste("Predicted Perpetrator Race:", predicted_race))
```

### Predict the perpetrator's sex
```{r predict_sex}
predicted_sex <- predict(sex_model, newdata = single_data_point)
print(paste("Predicted Perpetrator Sex:", predicted_sex))
```

### Look at predicted probabilities
Generate all combinations for a specific victim profile
```{r specific_combos}
specific_combinations <- expand.grid(
  VIC_RACE = "WHITE", 
  VIC_SEX = "M",       
  VIC_AGE_GROUP = "25-44", 
  PERP_RACE = unique(clean_data$PERP_RACE),
  PERP_SEX = unique(clean_data$PERP_SEX),
  PERP_AGE_GROUP = unique(clean_data$PERP_AGE_GROUP)
)
# Predict probabilities for the specific victim profile
predicted_probabilities_age_model <- predict(age_model, newdata = specific_combinations, type = "probs")

# Combine the predictions with the specific combinations
results_with_probabilities_age_model <- cbind(specific_combinations, predicted_probabilities_age_model)

# Predict probabilities for the specific victim profile
predicted_probabilities_race_model <- predict(race_model, newdata = specific_combinations, type = "probs")

# Combine the predictions with the specific combinations
results_with_probabilities_race_model <- cbind(specific_combinations, predicted_probabilities_race_model)

# Predict probabilities for the specific victim profile
predicted_probabilities_sex_model <- predict(sex_model, newdata = specific_combinations, type = "probs")

# Combine the predictions with the specific combinations
results_with_probabilities_sex_model <- cbind(specific_combinations, predicted_probabilities_sex_model)

# View the results
head(results_with_probabilities_age_model)
# View the results
head(results_with_probabilities_race_model)
# View the results
head(results_with_probabilities_sex_model)


# Initialize an empty data frame to store the results
results_combined <- data.frame(
  PERP_Sex = character(),
  PERP_Age_Group = character(),
  PERP_Race_Group = character(),
  Combined_Probability = numeric(),
  stringsAsFactors = FALSE
)


for (sex in unique_perp_sex) {
  for (age_group in unique_perp_age_group) {
    for (race_group in unique_perp_race) {
      
      # Extract the corresponding probabilities from each model
      age_prob <- results_with_probabilities_age_model[[age_group]][results_with_probabilities_age_model$PERP_SEX == sex & 
                                                                     results_with_probabilities_age_model$PERP_RACE == race_group]
      age_prob <- age_prob[1]
      
      race_prob <- results_with_probabilities_race_model[[race_group]][results_with_probabilities_race_model$PERP_SEX == sex & 
                                                                       results_with_probabilities_race_model$PERP_AGE_GROUP == age_group]
      race_prob <- race_prob[1]
      
      sex_prob <- results_with_probabilities_sex_model[[sex]][results_with_probabilities_sex_model$PERP_AGE_GROUP == age_group & 
                                                               results_with_probabilities_sex_model$PERP_RACE == race_group]
      sex_prob <- sex_prob[1]
      
      # Ensure there's only one value returned, if not handle accordingly
      if (length(age_prob) == 1 && length(race_prob) == 1 && length(sex_prob) == 1) {
        
        # Calculate the combined probability
        combined_prob <- age_prob * race_prob * sex_prob
        
        # Store the result
        results_combined <- rbind(results_combined, data.frame(PERP_Sex = sex, PERP_Age_Group = age_group, PERP_Race_Group = race_group, Combined_Probability = combined_prob))
      }
    }
  }
}

# View the results
# Sort the data frame by the Combined_Probability column in descending order
results_combined_sorted <- results_combined[order(-results_combined$Combined_Probability), ]
head(results_combined_sorted)



```
## All vic combinations
```{r all_vic_comb}
suppressWarnings({# Initialize an empty data frame to store the results
results_combined_ALL <- data.frame(
  VIC_Sex = character(),
  VIC_Age_Group = character(),
  VIC_Race_Group = character(),
  PERP_Sex = character(),
  PERP_Age_Group = character(),
  PERP_Race_Group = character(),
  Combined_Probability = numeric(),
  stringsAsFactors = FALSE
)
results_combined_ALL_filtered <- data.frame(
  VIC_Sex = character(),
  VIC_Age_Group = character(),
  VIC_Race_Group = character(),
  PERP_Sex = character(),
  PERP_Age_Group = character(),
  PERP_Race_Group = character(),
  Count = integer(),
  Probability = numeric(),
  stringsAsFactors = FALSE
)

# Loop through each combination of sex, age group, and race group
for (sex_vic in unique_vic_sex){
  for (age_group_vic in unique_vic_age_group){
    for (race_group_vic in unique_vic_race){
      specific_combinations <- expand.grid(
        VIC_RACE = race_group_vic, 
        VIC_SEX = sex_vic,       
        VIC_AGE_GROUP = age_group_vic, 
        PERP_RACE = unique(clean_data$PERP_RACE),
        PERP_SEX = unique(clean_data$PERP_SEX),
        PERP_AGE_GROUP = unique(clean_data$PERP_AGE_GROUP))
      # Predict probabilities for the specific victim profile
      predicted_probabilities_age_model <- predict(age_model, newdata = specific_combinations, type = "probs")
      
      # Combine the predictions with the specific combinations
      results_with_probabilities_age_model <- cbind(specific_combinations, predicted_probabilities_age_model)
      
      # Predict probabilities for the specific victim profile
      predicted_probabilities_race_model <- predict(race_model, newdata = specific_combinations, type = "probs")
      
      # Combine the predictions with the specific combinations
      results_with_probabilities_race_model <- cbind(specific_combinations, predicted_probabilities_race_model)
      
      # Predict probabilities for the specific victim profile
      predicted_probabilities_sex_model <- predict(sex_model, newdata = specific_combinations, type = "probs")
      
      # Combine the predictions with the specific combinations
      results_with_probabilities_sex_model <- cbind(specific_combinations, predicted_probabilities_sex_model)
      
      filtered_data <- clean_data %>%
        filter(VIC_RACE == race_group_vic, VIC_SEX == sex_vic, VIC_AGE_GROUP == age_group_vic)
      
      # Calculate the frequency of each perpetrator combination
      perp_combinations <- filtered_data %>%
        group_by(PERP_RACE, PERP_SEX, PERP_AGE_GROUP) %>%
        summarise(Count = n(), .groups = 'drop') %>%
        mutate(Probability = Count / sum(Count))  # Calculate observed probabilities
      
      # Ensure that the columns match between the two data frames before rbind
      perp_comb_binded <- data.frame(
        VIC_Sex = rep(sex_vic, nrow(perp_combinations)),
        VIC_Age_Group = rep(age_group_vic, nrow(perp_combinations)),
        VIC_Race_Group = rep(race_group_vic, nrow(perp_combinations)),
        PERP_Sex = perp_combinations$PERP_SEX,
        PERP_Age_Group = perp_combinations$PERP_AGE_GROUP,
        PERP_Race_Group = perp_combinations$PERP_RACE,
        Count = perp_combinations$Count,
        Probability = perp_combinations$Probability,
        stringsAsFactors = FALSE
      )
      
      # Store the result
      results_combined_ALL_filtered <- rbind(results_combined_ALL_filtered, perp_comb_binded)

      for (sex in unique_perp_sex) {
        for (age_group in unique_perp_age_group) {
          for (race_group in unique_perp_race) {
            
            # Extract the corresponding probabilities from each model
            age_prob <- results_with_probabilities_age_model[[age_group]][results_with_probabilities_age_model$PERP_SEX == sex & 
                                                                           results_with_probabilities_age_model$PERP_RACE == race_group]
            age_prob <- age_prob[1]
            
            race_prob <- results_with_probabilities_race_model[[race_group]][results_with_probabilities_race_model$PERP_SEX == sex & 
                                                                             results_with_probabilities_race_model$PERP_AGE_GROUP == age_group]
            race_prob <- race_prob[1]
            
            sex_prob <- results_with_probabilities_sex_model[[sex]][results_with_probabilities_sex_model$PERP_AGE_GROUP == age_group & 
                                                                     results_with_probabilities_sex_model$PERP_RACE == race_group]
            sex_prob <- sex_prob[1]
            
            # Ensure there's only one value returned, if not handle accordingly
            if (length(age_prob) == 1 && length(race_prob) == 1 && length(sex_prob) == 1) {
              
              # Calculate the combined probability
              combined_prob <- age_prob * race_prob * sex_prob
              
              # Store the result
              results_combined_ALL <- rbind(results_combined_ALL, data.frame(
                VIC_Sex = sex_vic, 
                VIC_Age_Group = age_group_vic, 
                VIC_Race_Group = race_group_vic, 
                PERP_Sex = sex, 
                PERP_Age_Group = age_group, 
                PERP_Race_Group = race_group, 
                Combined_Probability = combined_prob,
                stringsAsFactors = FALSE
              ))
            }
          }
        }
      }
    }
  }
}

# View the results
# Sort the data frame by the Combined_Probability column in descending order
results_combined_sorted_ALL <- results_combined_ALL[order(-results_combined_ALL$Combined_Probability), ]
head(results_combined_sorted_ALL)

results_combined_sorted_ALL_filtered <- results_combined_ALL_filtered[order(-results_combined_ALL_filtered$Probability), ]
head(results_combined_sorted_ALL_filtered)})

```

```{r stat_power}
# Set minimum count threshold, lets look at the power of all counts
min_count_threshold <- 1

# Filter the results by this threshold
results_significant <- results_combined_sorted_ALL_filtered %>%
  filter(Count >= min_count_threshold)

# View the filtered significant results
head(results_significant)


# Number of unique levels in each category (n)
n_race <- length(unique(results_significant$PERP_Race_Group))
n_sex <- length(unique(results_significant$PERP_Sex))
n_age_group <- length(unique(results_significant$PERP_Age_Group))

# Calculate degrees of freedom (DOF)
df <- (n_race - 1) * (n_sex - 1) * (n_age_group - 1)
calculate_power <- function(row,df) {
    effect_size <- 0.3 
    alpha <- 0.05 
    group_sample_size <- as.numeric(row['Count'])
    df_group <- df
    power_result <- pwr.chisq.test(w = effect_size, N = group_sample_size, df = df_group, sig.level = alpha)
    return(power_result$power)
}

results_significantwpower <- results_significant %>%
    rowwise() %>%
    mutate(Power = calculate_power(cur_data(),df))
head(results_significantwpower)
# A good power level cutoff is 80%.  This would mean that there is an 80% chance of detecting an effect if one truly exists.
# So lets look at values that are above the cutoff of 80%.
results_significantwpowergt80 <- results_significantwpower %>% filter(Power>0.8)
# This shows only 24 of the ~967 combinations have enough power to say there is this 80% chance.  We would need more data in the other groups to have the same power.

```
### Look at filtered_data on myself
We saw earlier the model predicted that I (MALE, 34 years old, and WHITE) would most likely be shot by someone who was UNKNOWN or BLACK, similar age and race.  However, we now can see how much power there is behind that prediction.  And it appears not much.  The probabilities aren't much different, but the power is less than 20%.  This means that we can't say there is a 20% chance of detecting an effect if one truly exists here.  
```{r filtered_data_on_self_w_power}
results_significantwpower[results_significantwpower$VIC_Sex=="M" & results_significantwpower$VIC_Age_Group=="25-44"& results_significantwpower$VIC_Race_Group=="WHITE",]
```
# Bias Identification/Conclusion  
  This dataset is strictly for New York City.  So the data found here might not apply outside of this city in other parts of America or the world at large.  Since I live in New York, but outside of New York City (source of this dataset), my **personal bias** might be that there isn't as many shootings.  However, if I were to spend time in New York City, my **personal bias** might be more corrected by what I am able to observe, **mitigating** its overall effect on my analysis/conclusions.  

  From this dataset, there is a significant amount of UNKNOWN data.  However, with the UNKNOWN data aside, it might appear that Black Males have the largest correlation with shooting.  Therefore, there could be bias built into this dataset to conclude UNKNOWN or Black Males are the number one perpetrators in shooting cases.  The power analysis I did showed this because only 14 of the ~967 combinations had a power level greater than 80% chance of detecting an effect if one truly exists.  Thus, the models I made would carry the bias of the data into them as they did.  We would need more data in the other missing groups to have more statistical power.  It could be that another race/gender/age has been getting away with shooting or murder much easier.  Therefore, garbage in, garbage out might apply here to the model's prediction.  
  
  When we look at the models fit to the data we observe for a victim of my age, gender, and race [34,M,White] that the perpetrator is most likely a Male, between 25-44, and of an UNKNOWN race.  Excluding the UNKNOWN, the perpetrator is most likely a Male, between 25-44, and Black with less than a 7% chance of this combination.  And interesting enough, at only a 6% chance, the same categories as myself would predict a perpetrator.  This tells me that there is a positive correlation between the victim and the perpetrator in terms of their race, sex, and age.  We further investigated the statistical power of these predictions of the model's by looking at the filtered data without the model and performing a statistical power test.  We found that the probabilities aren't much different, but the statistical power is less than 20%.  This means that we can't say there is a 20% chance of detecting an effect if one truly exists here.  
  
  In conclusion, all we know is that there is a significant amount of UNKNOWN shooters out there, and from what we do know, a large number of them appear to be Black Males from this dataset.  Using a model based on this dataset that predicts the perpetrator given the victim's race, age, and gender likely will give very biased results as shown in this analysis because of this missing data and also because of the lack of diverse data.  Hopefully with time, we can solve more of the mystery cases out there to fill in the missing data.  With time, there might be more data that adds to statistical significance of potential models or analysis.  

